Sequence Labeling for Gait Analysis using LSTM

### 2015-05-09 paper/paper.tex
Fix typos

The Foot Velocity Algorithm (FVA) proposed by
\cite{OConnor2007}
belongs to a category of algorithms that use heuristics such as the velocity
and acceleration of heel and toe markers to detect motion events.
There are other examples of such algorithms, notably the one developped by
\cite{Hreljac2000}.
As these algorithms are quite similar, we will restrict the discussion
to the FVA.

@@

% This makes it a popular choice in practice.
The FVA is widely used in practice.
The reason for algorithm's popularity lies in the ease with which it can be
implemented:
it does not require any data preprocessing beyond simple signal processing and
it does not have to be trained.
That said, the FVA has several problems.
First, it makes questionable assumptions about the relationship between the
marker location and the gait cycle.
For instance, it is not clear that the marker velocity peaks exactly coincide
with the events to be identified.
Secondly, the algorithm is very sensitive to a threshold that has to be applied
in order to restrict the search for heel strike candidates.
The value of this threshold has to be manually tuned for each subject.
If not chosen carefully, the algorithm fails catastrophically.
Finally, the accuracy of FVA is bad when compared to more sophisticated methods
such as neural network and LSTM
(cf. Section~\ref{sec:Results}).

@@

\cite{Miller2009}
treats event detection as a classification problem and uses a classical
(feed-forward) neural network to perform the classification.
Each sample is classified individually based on a input window centered around
the desired sample.
At every time step, the window center is advanced by one sample
(so-called sliding window technique).
The use of a window is necessary to take into account the temporal dependencies
that exists between neighboring samples.%
\footnote
{
    Ideally, the window size should not fixed vary as a function of the forward 
    motion of the foot.
    The reason for this is that a fixed size window cannot guarantee that every
    window contained a sufficient amount of context because of the varying foot
    speed.
    For simplicity, we decided to use a large fixed window size in our
    implementation.
    % in our implementation because
    % we found it difficult to reliably calculate the foot displacement from our
    % data.
}

The input is not the actual motion capture data, but rather position,
instantaneous velocity and instantaneous acceleration features extracted from
that data.
Furthermore, PCA is used to reduce the number of input variables from several
hundred to 50.
This dimensionality reduction is necessary because the total input size grows
linearly with the window size and the number of markers.

>>>>

The Foot Velocity Algorithm (FVA)
\cite{OConnor2007}
belongs to a category of algorithms that use heuristics such as the velocity
and acceleration of heel and toe markers to detect motion events.
There are other examples of such algorithms, notably the one by Hreljac and 
Marshall
\cite{Hreljac2000}.
As these algorithms are quite similar, we will restrict our discussion
to the FVA.

@@

% This makes it a popular choice in practice.
The FVA is widely used in practice.
The reason for the algorithm's popularity lies in the ease with which it can be
implemented:
it does not require any data preprocessing beyond some simple signal processing 
and it does not have to be trained.
That said, the FVA has several problems.
First, it makes questionable assumptions about the relationship between the
marker location and the gait cycle.
For instance, it is not clear that the marker velocity peaks exactly coincide
with the events to be identified.
Secondly, the algorithm is very sensitive to a threshold that has to be applied
in order to restrict the search for heel strike candidates.
The value of this threshold has to be manually tuned for each subject.
If not chosen carefully, the algorithm fails catastrophically.
Finally, the accuracy of FVA is bad when compared to more sophisticated methods
such as neural network and LSTM
(cf. Section~\ref{sec:Results}).

@@

\cite{Miller2009}
treats event detection as a classification problem and uses a classical
(feed-forward) neural network to perform the classification.
Each sample is classified individually based on a input window centered around
the desired sample.
At every time step, the window center is advanced by one sample
(so-called sliding window technique).
The use of a window allows to take into account the temporal dependencies
that exists between neighboring samples.%
\footnote
{
    Ideally, the window size should not be fixed but vary as a function of the 
    forward motion of the foot.
    The reason for this is that a fixed size window cannot guarantee that every
    window contains a sufficient amount of context because of the varying foot
    speed.
    For simplicity, we decided to use a fixed window size in our implementation
    of this method.
}

The input is not the actual motion capture data but rather position,
instantaneous velocity and instantaneous acceleration features extracted from
that data.
Furthermore, PCA is used to reduce the number of input variables from several
hundred to 50.
This dimensionality reduction is necessary because the number of inputs grows
linearly with the window size and the number of markers.
